Supplementary material for ``Screening Sinkhorn Algorithm for regularized Optimal Transport''

### 2019-05-16 tex/v3/section_3_algorithm.tex
new

Pseudocode of our proposed algorithm is shown in Algorithm~\ref{screenkhorn}. %~\cite{nocedal1980,nocedal2006numerical}
It is worth to note that Screenkhorn algorithm uses only the restricted parts $K_{(I_{\varepsilon,\kappa}, J_{\varepsilon,\kappa})},$ $K_{(I_{\varepsilon,\kappa}, J^\complement_{\varepsilon,\kappa})},$ and $K_{(I^\complement_{\varepsilon,\kappa}, J_{\varepsilon,\kappa})}$ of the Gibbs kernel $K$ for calculating the objective function $\Psi_{\varepsilon, \kappa}$ and its gradients, in contrast to Sinkhorn algorithm which performs alternating updates of all rows and columns of $K.$
Furthermore, it consists of two steps: the first one is an initialization providing the active sets $I_{\varepsilon,\kappa}$, $J_{\varepsilon,\kappa}$. 
The second is a constrained L-BFGS-B for the stacked variable $\theta=(u_{I_{\varepsilon,\kappa}},v_{J_{\varepsilon,\kappa}}).$ 
The L-BFGS-B solver handles box constraints on variables, and it becomes more efficient when these box bounds are carefully determined for problem~\eqref{screen-sinkhorn_second_def}. 
The following lemma (proof in supplementary material) expresses these bounds that are pre-calculated in the initialization step of Screenkhorn.

\begin{lemma}
\label{lemma_bounds_of_usc_and_vsc}
Let $(u^{\text{sc}}, v^{\text{sc}})$ be an optimal solution of problem~\eqref{screen-sinkhorn_second_def}, and $K_{\min} = \min\limits_{i\in I_{\varepsilon,\kappa},j \in J_{\varepsilon,\kappa}}K_{ij}$. Then,
one has
\begin{equation}
\label{bound_on_u}
\frac \varepsilon\kappa \vee \frac{\min_{i \in I_{\varepsilon,\kappa}}\mu_i}{\varepsilon (m- m_b) + \varepsilon \vee \frac{\max_{j\in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon\kappa K_{\min}} m_b} \leq e^{u^{\text{sc}}_i} \leq \frac \varepsilon\kappa\vee \frac{\max_{i \in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min}},
\end{equation}
and
\begin{equation}
\label{bound_on_v}
\varepsilon\kappa \vee \frac{\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\varepsilon(n- n_b) + \varepsilon \vee \frac{\kappa\max_{i\in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min} } n_b} \leq e^{v^{\text{sc}}_j} \leq \varepsilon\kappa \vee \frac{\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min} }
\end{equation}
for all $i\in I_{\varepsilon,\kappa}$ and $j\in J_{\varepsilon,\kappa}$.
\end{lemma}

>>>>

Pseudocode of our proposed algorithm is shown in Algorithm~\ref{screenkhorn}. %~\cite{nocedal1980,nocedal2006numerical}
It is worth to note that Screenkhorn algorithm uses only the restricted parts $K_{(I_{\varepsilon,\kappa}, J_{\varepsilon,\kappa})},$ $K_{(I_{\varepsilon,\kappa}, J^\complement_{\varepsilon,\kappa})},$ and $K_{(I^\complement_{\varepsilon,\kappa}, J_{\varepsilon,\kappa})}$ of the Gibbs kernel $K$ for calculating the objective function $\Psi_{\varepsilon, \kappa}$ and its gradients, in contrast to Sinkhorn algorithm which performs alternating updates of all rows and columns of $K.$
Furthermore, it consists of two steps: the first one is an initialization providing the active sets $I_{\varepsilon,\kappa}$, $J_{\varepsilon,\kappa}$. 
The second is a constrained L-BFGS-B for the stacked variable $\theta=(u_{I_{\varepsilon,\kappa}},v_{J_{\varepsilon,\kappa}}).$ 
The L-BFGS-B solver handles box constraints on variables, and it becomes more efficient when these box bounds are carefully determined for problem~\eqref{screen-sinkhorn_second_def}. 
The following proposition (proof in supplementary material) expresses these bounds that are pre-calculated in the initialization step of Screenkhorn.

\begin{proposition}
\label{prop:bounds_of_usc_and_vsc}
Let $(u^{\text{sc}}, v^{\text{sc}})$ be an optimal solution of problem~\eqref{screen-sinkhorn_second_def}, and $K_{\min} = \min\limits_{i\in I_{\varepsilon,\kappa},j \in J_{\varepsilon,\kappa}}K_{ij}$. Then,
one has
\begin{equation}
\label{bound_on_u}
\frac \varepsilon\kappa \vee \frac{\min_{i \in I_{\varepsilon,\kappa}}\mu_i}{\varepsilon (m- m_b) + \varepsilon \vee \frac{\max_{j\in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon\kappa K_{\min}} m_b} \leq e^{u^{\text{sc}}_i} \leq \frac \varepsilon\kappa\vee \frac{\max_{i \in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min}},
\end{equation}
and
\begin{equation}
\label{bound_on_v}
\varepsilon\kappa \vee \frac{\min_{j \in J_{\varepsilon,\kappa}}\nu_j}{\varepsilon(n- n_b) + \varepsilon \vee \frac{\kappa\max_{i\in I_{\varepsilon,\kappa}} \mu_i}{m\varepsilon K_{\min} } n_b} \leq e^{v^{\text{sc}}_j} \leq \varepsilon\kappa \vee \frac{\max_{j \in J_{\varepsilon,\kappa}} \nu_j}{n\varepsilon K_{\min} }
\end{equation}
for all $i\in I_{\varepsilon,\kappa}$ and $j\in J_{\varepsilon,\kappa}$.
\end{proposition}
